---
title: "Measuring_gentrification"
author: "Regy Septian and Xiaxin Tang"
date: "2024-05-01"
output: html_document
---
gentrification measure:
Dependent Variable: gentrification score:
1. change of percentage of high income B19013_001E
2. change of ethnic composition B02001_003
3. change of high education group B23006_001
4. change of percentage of professional employment B23025_001
5. change of percentage of housing above median rent 	
B25031_001
6. change of income statues
7. change of median age B01002_001E

Independent Variables:
1. Change of Public transportation
2. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

install.packages("riem")
install.packages("tidyverse")
installed.packages("sf")
installed.packages("lubridate")
install.packages("tidycensus")
library(tidycensus)
library(tidyverse)
library(sf)
library(lubridate)
library(tigris)
library(tidycensus)
library(viridis)
library(riem)
library(gridExtra)
library(knitr)
library(kableExtra)
library(RSocrata)

tidycensus::census_api_key("7c984127f786a51716ed4f5115e1b08c3d2033f1", overwrite = TRUE)

```

```{r get_census, message=FALSE, warning=FALSE, cache=TRUE, results = 'hide'}

install.packages("sf")
library(sf)
acs_variable_list.2009 <- load_variables(2009, #year
                                         "acs5", #five year ACS estimates
                                         cache = TRUE)

PHLCensus_2019 <- 
  get_acs(geography = "tract", 
          variables = c("B01003_001E", "B19013_001E", 
                        "B01002_001E", "B02001_002E",
                        "B23006_023E", "B08301_001E", "B08301_010E"),
          year = 2019, 
          state = "PA", 
          county = "Philadelphia",
          geometry = TRUE, 
          output = "wide") %>%
  rename(Total_Pop_t2 =  B01003_001E,
         Med_Inc_t2 = B19013_001E,
         Med_Age_t2 = B01002_001E,
         Black_Pop_t2 = B02001_002E,
         Education_attainment_t2 = B23006_023E,
         Median_rent_t2 = B08301_001E,
         Total_Public_Trans_t2 = B08301_010E) %>%
  select(Total_Pop_t2, Med_Inc_t2, Black_Pop_t2, Education_attainment_t2, 
         Total_Public_Trans_t2, Med_Age_t2, Median_rent_t2, GEOID) %>%
  mutate(Percent_black_t2 = Black_Pop_t2 / Total_Pop_t2) %>% 
  st_transform(crs = "EPSG:4326") %>% 
  st_drop_geometry()

PHLCensus_2014 <- 
  get_acs(geography = "tract", 
          variables = c("B01003_001E", "B19013_001E", 
                        "B01002_001E", "B02001_002E",
                        "B23006_023E", "B08301_001E", "B08301_010E"),
          year = 2014, 
          state = "PA", 
          county = "Philadelphia",
          geometry = TRUE, 
          output = "wide") %>%
  rename(Total_Pop_t1 =  B01003_001E,
         Med_Inc_t1 = B19013_001E,
         Med_Age_t1 = B01002_001E,
         Black_Pop_t1 = B02001_002E,
         Education_attainment_t1 = B23006_023E,
         Median_rent_t1 = B08301_001E,
         Total_Public_Trans_t1 = B08301_010E) %>%
  select(Total_Pop_t1, Med_Inc_t1, Black_Pop_t1, Education_attainment_t1,
         Total_Public_Trans_t1, Med_Age_t1, Median_rent_t1, 
         GEOID, geometry) %>%
  mutate(Percent_black_t1 = Black_Pop_t1 / Total_Pop_t1) %>% 
  st_transform(crs = "EPSG:4326")

PHLCensus <- PHLCensus_2019 %>% 
  left_join(., PHLCensus_2014)

```



```{r cars}
PHLCensus <- PHLCensus %>% 
  mutate(change_pop = (Total_Pop_t2 - Total_Pop_t1)/Total_Pop_t1,
         change_black_pop = (Black_Pop_t2 - Black_Pop_t1)/Black_Pop_t1,
         change_high_income = (Med_Inc_t2 - Med_Inc_t1)/Med_Inc_t1,
         change_high_edu = (Education_attainment_t2 - Education_attainment_t1)/Education_attainment_t1,
         change_public_tran = (Total_Public_Trans_t2 - Total_Public_Trans_t1)/mean(Total_Public_Trans_t1),
         change_age = (Med_Age_t2 - Med_Age_t1)/Med_Age_t1,
         change_rent = (Median_rent_t2 - Median_rent_t1)/Median_rent_t1,
         change_perc_black = (Percent_black_t2 - Percent_black_t1)/Percent_black_t1 * -1,
         gent_score = change_perc_black + change_rent - change_age +
           change_high_edu + change_high_income) %>% 
  mutate(cate_gent = ifelse(gent_score < 0, "degentrified",         
                    ifelse(gent_score > 0 & gent_score < 0.61, "slightly gentrified",
                    ifelse(gent_score > 0.61 & gent_score < 1.345, "gentrified", "highly gentrified"))))


PHLCensus <- PHLCensus %>% 
  mutate(change_pop = (Total_Pop_t2 - Total_Pop_t1)/Total_Pop_t1,
         change_black_pop = (Black_Pop_t2 - Black_Pop_t1)/Black_Pop_t1,
         change_high_income = (Med_Inc_t2 - Med_Inc_t1)/Med_Inc_t1,
         change_high_edu = (Education_attainment_t2 - Education_attainment_t1)/Education_attainment_t1,
         change_public_tran = (Total_Public_Trans_t2 - Total_Public_Trans_t1)/mean(Total_Public_Trans_t1),
         change_age = (Med_Age_t2 - Med_Age_t1)/Med_Age_t1,
         change_rent = (Median_rent_t2 - Median_rent_t1)/Median_rent_t1,
         change_perc_black = (Percent_black_t2 - Percent_black_t1)/Percent_black_t1 * -1,
         gent_score = change_rent + change_high_income) %>% 
  mutate(cate_gent = ifelse(gent_score < 0, "degentrified",         
                    ifelse(gent_score > 0 & gent_score < 0.61, "slightly gentrified",
                    ifelse(gent_score > 0.61 & gent_score < 1.345, "gentrified", "highly gentrified"))))

quartiles <- quantile(PHLCensus$gent_score, probs = c(0.25, 0.5, 0.75))
quartiles

PHLCensus$Black_Pop_t2[PHLCensus$Black_Pop_t2 == 0] <- 2

PHLCensus$Black_Pop_t1[PHLCensus$Black_Pop_t1 == 0] <- 1

PHLCensus$gent_score[is.infinite(PHLCensus$gent_score)] <- 0

PHLCensus$gent_score[is.infinite(PHLCensus$gent_score)] <- 0

PHLCensus$gent_score[is.na(PHLCensus$gent_score)] <- 0

PHLCensus <- st_as_sf(PHLCensus, crs = 4326)

ggplot(data = PHLCensus, aes(x = gent_score)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Random Data", x = "Value", y = "Frequency")

```



Parks

```{r pressure, echo=FALSE}
install.packages("mapview")
library(mapview)
Parks_2019 <- st_read("PPR_Properties.geojson") %>% 
  select(PUBLIC_NAME) %>% 
  st_transform(crs = "EPSG:4326")

# Calculate centroids
centroids_2019 <- st_centroid(Parks_2019)


Parks_2016 <- st_read("PhillyFPC_PPR_parks_point.shp") %>% 
  select("NAME")  %>% 
  st_transform(crs = "EPSG:4326")

mapview(PHLCensus)
mapview(Parks_2019)

PHLCensus_park_2016 <- PHLCensus %>%
  st_join(., Parks_2016) %>%
  group_by(GEOID) %>%
  mutate(total_parks_t1 = n(),
         total_parks_t1 = case_when(
           is.na(NAME) ~ 0,
           TRUE ~ total_parks_t1
         )) %>% 
  distinct(GEOID, total_parks_t1)



PHLCensus_park_2019 <- PHLCensus %>%
  st_join(., centroids_2019) %>%
  group_by(GEOID) %>%
  mutate(total_parks_t2 = n(),
         total_parks_t2 = case_when(
           is.na(PUBLIC_NAME) ~ 0,
           TRUE ~ total_parks_t2
         )) %>% 
  distinct(GEOID, total_parks_t2)

PHLCensus <- PHLCensus %>% 
  left_join(., PHLCensus_park_2016) %>% 
  left_join(., PHLCensus_park_2019)



```
Buses

```{r pressure, echo=FALSE}

bus_2014 <- read_csv("Bus_Ridership_by_Census_Tract_2014.csv") %>% 
  filter(Season == "Fall 2014") %>% 
  group_by(Census_Tract_ID) %>% 
  summarise(Total_On_t1 = sum(On_)) %>% 
  select("Total_On_t1", "Census_Tract_ID")%>% 
  rename(GEOID = Census_Tract_ID) 

bus_2014$GEOID <- as.character(bus_2014$GEOID)

bus_2019 <- read_csv("Bus_Ridership_by_Census_Tract_2019.csv") %>% 
  filter(Season == "Fall 2019") %>% 
  group_by(Census_Tract_ID) %>% 
  summarise(Total_On_t2 = sum(On_)) %>% 
  select("Total_On_t2", "Census_Tract_ID") %>% 
  rename(GEOID = Census_Tract_ID)
bus_2019$GEOID <- as.character(bus_2019$GEOID)

PHLCensus_Bus_2014 <- PHLCensus %>% 
  left_join(., bus_2014, by = c("GEOID"))

PHLCensus <- PHLCensus_Bus_2014 %>% 
  left_join(., bus_2019, by = c("GEOID"))



```


Permits

```{r pressure, echo=FALSE}
Permits2 <- st_read("permits_2016-present/permits.shp") %>% 
  filter(grepl("^2019", permitissu)) %>% 
  select("objectid")

permits_per_tract2 <- PHLCensus %>% 
  st_join(., Permits2) %>% 
  group_by(GEOID) %>% 
  summarise(num_permits_t2 = n()) %>% 
  st_drop_geometry()

PHLCensus <- PHLCensus %>% 
  left_join(., permits_per_tract2)

Permits1 <- st_read("permits_2007-2015/permits.shp") %>% 
  filter(grepl("^2014", permitissu)) %>% 
  select("objectid")

permits_per_tract1 <- PHLCensus %>% 
  st_join(., Permits1) %>% 
  group_by(GEOID) %>% 
  summarise(num_permits_t1 = n()) %>% 
  st_drop_geometry()

PHLCensus <- PHLCensus %>% 
  left_join(., permits_per_tract1)



```
Food Access

```{r pressure, echo=FALSE}
food_2019 <- st_read("Food_access/NeighborhoodFoodRetail.geojson") %>% 
  mutate(HPSS_ACCESS_t2 = ifelse(HPSS_ACCESS == "No Access", 0, 
  ifelse(HPSS_ACCESS == "Low Access", 1, 2))) %>% 
  select("HPSS_ACCESS_t2")

# Calculate mean of column A (ignoring NAs)
mean_A <- mean(food_2019$HPSS_ACCESS_t2, na.rm = TRUE)

# Replace NA values in column A with the mean
food_2019$HPSS_ACCESS_t2[is.na(food_2019$HPSS_ACCESS_t2)] <- mean_A

food_2019 <- st_centroid(food_2019)

Access_census_2019 <- PHLCensus %>% 
  st_join(., food_2019) %>% 
  group_by(GEOID) %>% 
  summarise(Access_t2 = mean(HPSS_ACCESS_t2)) %>% 
  st_drop_geometry()
  

food_2014 <- st_read("Food_access/Walkable_Access_Healthy_Food.geojson") %>% 
  mutate(HPSS_ACCESS_t1 = ifelse(ACCESS_ == "No Access", 0,         
  ifelse(ACCESS_ == "Low Access", 1,
  ifelse(ACCESS_ == "Moderate Access", 2, 3)))) %>% 
  select("HPSS_ACCESS_t1")

food_2014 <- st_make_valid(food_2014)

mapview(food_2014)

food_2014 <- st_centroid(food_2014)
# Calculate mean of column A (ignoring NAs)
mean_B <- mean(food_2019$HPSS_ACCESS_t2, na.rm = TRUE)

# Replace NA values in column A with the mean
food_2014$HPSS_ACCESS_t1[is.na(food_2014$HPSS_ACCESS_t1)] <- mean_B


Access_census_2014 <- PHLCensus %>% 
  st_join(., food_2014) %>% 
  group_by(GEOID) %>% 
  summarise(Access_t1 = mean(HPSS_ACCESS_t1)) %>% 
  st_drop_geometry()

PHLCensus <- PHLCensus %>% 
  left_join(., Access_census_2014) %>% 
  left_join(., Access_census_2019) 


```

School

```{r pressure, echo=FALSE}
# Check the CRS of PHLCensus
PHLCensus_crs <- st_crs(PHLCensus)

school_2019 <- st_read("Schools.geojson") %>% 
  select("OBJECTID")

school_2019 <- st_transform(school_2019, crs = PHLCensus_crs)

school_count_2019 <- PHLCensus %>% 
  st_join(., school_2019) %>% 
  group_by(GEOID) %>% 
  summarise(school_count_2019 = n()) %>% 
  st_drop_geometry()

school_2014 <- st_read("PhiladelphiaSchool_Facilities2016/PhiladelphiaSchool_Facilities2016.shp") %>% 
  select("OBJECTID")

school_2014 <- st_transform(school_2014, crs = PHLCensus_crs)

school_count_2014 <- PHLCensus %>% 
  st_join(., school_2014) %>% 
  group_by(GEOID) %>% 
  summarise(school_count_2014 = n()) %>% 
  st_drop_geometry()

PHLCensus <- PHLCensus %>% 
  left_join(., school_count_2014) %>% 
  left_join(., school_count_2019) 

```

crime
```{r pressure, echo=FALSE}

crime_2014 <- st_read("crime2014/incidents_part1_part2.shp") %>% 
  select("objectid")

crime_2014 <- st_transform(crime_2014, crs = PHLCensus_crs)

crime_2019 <- st_read("crime2019/incidents_part1_part2.shp") %>% 
  select("objectid")

crime_2019 <- st_transform(crime_2019, crs = PHLCensus_crs)



crime_count_2014 <- PHLCensus %>% 
  st_join(., crime_2014) %>% 
  group_by(GEOID) %>% 
  summarise(crime_count_2014 = n()) %>% 
  st_drop_geometry()

crime_count_2019 <- PHLCensus %>% 
  st_join(., crime_2019) %>% 
  group_by(GEOID) %>% 
  summarise(crime_count_2019 = n()) %>% 
  st_drop_geometry()

PHLCensus <- PHLCensus %>% 
  left_join(., crime_count_2014) %>% 
  left_join(., crime_count_2019) 

```


indego

```{r pressure, echo=FALSE}

indego_2016 <- read_csv("indego_2016/Indego_Trips_2016Q1.csv")
%>% 
  select("objectid")


indego_2019 <- read_csv("indego_2019/indego-trips-2019-q4.csv")
%>% 
  select("objectid")


```

simple regression

```{r pressure, echo=FALSE}
install.packages("zoo")
library(zoo)

PHLCensus <- PHLCensus %>% 
  mutate(park_change = (total_parks_t2 - total_parks_t1)/total_parks_t1,
         bus_change = (Total_On_t2 - Total_On_t1)/Total_On_t1,
         permits_change = (num_permits_t2 - num_permits_t1)/num_permits_t1,
         food_change = (Access_t2 - Access_t1)/Access_t1,
         school_change = (school_count_2019 - school_count_2014)/school_count_2014,
         crime_change = (crime_count_2019 - crime_count_2014)/crime_count_2014)

PHLCensus <- PHLCensus[complete.cases(PHLCensus$gent_score), ]

PHLCensus <- transform(PHLCensus, 
                       permits_change = ifelse(is.na(permits_change), mean(permits_change, na.rm = TRUE), permits_change),
                       food_change = ifelse(is.na(food_change), mean(food_change, na.rm = TRUE), food_change),
                       school_change = ifelse(is.na(school_change), mean(school_change, na.rm = TRUE), school_change),
                       crime_change = ifelse(is.na(crime_change), mean(crime_change, na.rm = TRUE), crime_change),
                       change_high_income = ifelse(is.na(change_high_income), mean(change_high_income, na.rm = TRUE), change_high_income),
                       bus_change = ifelse(is.na(bus_change),
mean(bus_change, na.rm = TRUE), bus_change))

PHLCensus$bus_change[is.infinite(PHLCensus$bus_change)] <- 0

PHLCensus <- PHLCensus %>% 
  mutate(gent_score2 = bus_change + permits_change + food_change + crime_change)

PHLCensus$gent_score2[is.infinite(PHLCensus$gent_score2)] <- 0

PHLCensus$change_black_pop[is.infinite(PHLCensus$change_black_pop)] <- 0

model2 <- lm(gent_score2 ~ permits_change + school_change + crime_change + change_high_income + change_high_edu + change_black_pop + change_pop, data = PHLCensus)

#PHLCensus <- PHLCensus %>% 
#  filter(cate_gent == "gentrified" | cate_gent == "highly gentrified")
# Fit the linear regression model


model <- lm(gent_score2 ~ change_rent + change_high_income + change_high_edu + change_black_pop , data = PHLCensus)


model.2 <- lm(change_rent ~ park_change + bus_change +
              permits_change + school_change + crime_change, data = PHLCensus)
summary(model)

ggplot(data = PHLCensus, aes(x = gent_score)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Random Data", x = "Value", y = "Frequency")

ggplot(data = PHLCensus, aes(x = permits_change)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Random Data", x = "Value", y = "Frequency")

ggplot(data = PHLCensus, aes(x = bus_change)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Random Data", x = "Value", y = "Frequency")

ggplot(data = PHLCensus, aes(x = change_age)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Histogram of Random Data", x = "Value", y = "Frequency")

```

